---
title: "Analysis - Daims Actinobacteria ENR4 assembly"
author: "Mads Albertsen"
date: "November 24, 2015"
output: html_document
---

## Introduction
This report documents the binning of a Actinobacteria genome from an enrichment reactor (ENR4). See [Daims et al., 2015: Complete Nitrification by Nitrospira Bacteria](https://doi.org/10.1038/nature16461) for further details.

## Load the mmgenome package
In case you haven't installed the [mmgenome package](http://madsalbertsen.github.io/mmgenome/), see the [Load data example](http://madsalbertsen.github.io/mmgenome/Load_data.html).
```{r Load_package, message=FALSE, warning=FALSE, results='hide'}
library("mmgenome")
```

## Import data
The Rmarkdown file [Load_data.Rmd](http://madsalbertsen.github.io/mmgenome/Load_data_ENR4.html) describes the loading of the data and can be imported using the `mmimport` function. However, the preprocessed data can also be downloaded directly from figshare: [Daims_ENR4](http://dx.doi.org/10.6084/m9.figshare.1613277). Hence, here we import the prepocessed data from figshare instead.
```{r Load_data, message=FALSE, warning=FALSE, results='hide'}
load("Daims_ENR4.RData")
```

## Data overview
The object `d` contains information on scaffolds and essential genes within the scaffolds. For each scaffold the dataset contains the following information: The columns `ENR4A`, `ENR4E`, `ENR4F` and `ENR6` contain the coverage information from 4 different samples; `PC1`, `PC2` and `PC3` contain coordinates of the three first principal components from a PCA analysis on tetranucleotide frequencies; `essential` contain information taxonomic information for each scaffold based on classification on essential genes; `rRNA` contain taxonomic information on scaffolds that have an associated 16S rRNA gene.
```{r overview_names}
colnames(d$scaffolds)
```

The basic statistics of the full dataset can be summarised using the `mmstats` function. 
```{r overview_stats}
mmstats(d, ncov = 4)
```

## Extract the Actinobacteria
The combination of the coverage of sample `ENR4A` and `ENR4E` provides the cleanest separation of the genomes and are used for binning. A subspace is defined that clearly seperates the Actinobacteria from the three other species.
```{r zoomA, fig.width=12, fig.height=10, fig.align='center'}
p <- mmplot(data = d, x = "ENR4A", y = "ENR4E", log.x = T, log.y = T, color = "essential") 

#p
#sel <- mmplot_locator(p)

sel <- data.frame(ENR4A  =  c(2.54, 3.72, 6.53, 5.86, 2.73),
                  ENR4E  =  c(5.45, 8.06, 6.63, 3.93, 3.93))

mmplot_selection(p, sel)
```

```{r extractA}
dA <- mmextract(d, sel)
```

```{r statsA}
mmstats(dA, ncov = 4)
```

## Using paried-end connections
Until now we have just used coverage profiles to extract scaffolds related to our genome of interest. However, some scaffolds might be present in many copies (repeats) and hence have a much higher coverage than the rest of the genome. In addition, some scaffolds will by chance have a slightly different coverage profile than the rest of the genome and thereby also been missed.

The function `mmplot_network` can be used to generate a network plot of scaffolds connected by paired-end reads. We start by plotting the scaffolds we have in our current subset.

```{r networkB, fig.width=10, fig.height=10, fig.align='center'}
mmplot_network(data = dA, network = pe, nconnections = 2, color = "essential", scale.links = 0.5)
```

All scaffolds seem to be nicely connected and we should be able to improve the assembly substantial in the refinement stage. To include repeats and other missed scaffolds we simply extract all scaffolds that are directly connected by paired-end reads to our current subset dB using mmextract_network. Only scaffolds directly connected to the subset is extracted.
```{r network_extractC}
dB <- mmextract_network(subset = dA, original = d, network = pe, nconnections = 5, type = "direct")
```

… and then plot the new subset. It seems like we included a single high coverage scaffold that was missed initially.
```{r networkC, fig.width=10, fig.height=10, fig.align='center'}
mmplot_network(data = dB, network = pe, nconnections = 1, color = "ENR4A", scale.links = 0.5)
```

```{r statsD}
mmstats(dB, ncov = 4)
```

## Export the scaffolds 
Now that we are happy with the genome bin, the scaffolds can be exported to a separate fasta file using `mmexport`. The genome were then re-assembled using the reads in the genome bin using the SPAdes assembler.
```{r mmexport}
mmexport(data=dB, assembly=assembly, file = "Actinobacteria.fa")
```
