---
title: "Analysis - Daims Nitrospira 2 GWW assembly"
author: "Mads Albertsen"
date: "November 24, 2015"
output: html_document
---
## Introduction
This report documents the binning of a Nitrospira genome from the GWW sample. See [Daims et al., 2015: Complete Nitrification by Nitrospira Bacteria](https://doi.org/10.1038/nature16461) for further details.

## Load the mmgenome package
The metagenome data is analysed using the [mmgenome package](http://madsalbertsen.github.io/mmgenome/).
```{r Load_package, message=FALSE, warning=FALSE, results='hide'}
library("mmgenome")
```

## Import data
The Rmarkdown file [Load_data.Rmd](http://madsalbertsen.github.io/mmgenome/Daims_GWW_Load_data.html) describes the loading of the data and can be imported using the `mmimport` function. However, the preprocessed data can also be downloaded directly from figshare: [Daims_GWW](http://dx.doi.org/10.6084/m9.figshare.1613284). Hence, here we import the prepocessed data from figshare instead.
```{r Load_data, message=FALSE, warning=FALSE, results='hide'}
load("Daims_GWW.RData")
```

## Extract Nitrospira 2
The second Nitrospira bin can also be relatively easy seperated from the other genomes.
```{r zoomA, fig.width=12, fig.height=10, fig.align='center', warning=FALSE, message=FALSE}
p <- mmplot(data = d, 
            x = "HPD", 
            y = "HPF1", 
            log.x = F, 
            log.y = F, 
            color = "essential", 
            minlength = 1000,
            factor.shape = "solid") +
  xlim(0, 50) +
  ylim(50, 600)

#p
#sel <- mmplot_locator(p)

sel <- data.frame(HPD  =  c(-0.483, 0.487, 12.9, 25.5, 23.4, 11.7, 2.75),
                  HPF1  =  c(210, 292, 307, 292, 209, 148, 148))

mmplot_selection(p, sel)
```

The scaffolds included in the defined subspace are extracted using the `mmextract` function. 
```{r extractA}
dA <- mmextract(d, sel)
```

The `mmstats` function applies to any extracted object. Hence, it can be used directly on the subset.
```{r statsA}
mmstats(dA, ncov = 2)
```

## Export the scaffolds
Finally the binned scaffolds are exported.
```{r export}
mmexport(data = dA, assembly = assembly, file = "GWW_Nitrospira2.fa")
```