General overview
=======================================

## Load data

```{r Loaddata, message=FALSE, warning=FALSE, results='hide'}
library("mmgenome")
knit('Load_data.Rmd', tangle=TRUE)
source(file="Load_data.R")
```

## Initial overview of the data
The function `calc.genome.stats` can be used to calculate basic statistics on the full dataset.
```{r Overview - stats}
sstats(d,e)
```
`tot.ess` is the total number of essential genes identified, where `uni.ess` is the number of unique essential genes. 

For more pretty plots we subset the data.
```{r subset}
ds <- subset(d, length > 3000)
```

## Differential coverage plots

The differential coverage plot is the starting point for most genome extractions.
```{r S1_S2_phylum_plot, fig.height=8, fig.width=10, warning=F}
ggplot(ds, aes(x = S1, y = S2, size = length, color = phylum)) + 
  scale_x_log10() + 
  scale_y_log10() + 
  geom_point(alpha=0.1, color = 'black') +
  geom_point(data=subset(ds, phylum != "NA"), shape = 1, alpha = 0.7) +
  scale_size_area(name= "Scaffold length", max_size=20) +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size = 5, shape = 19)))
```

## GC coverage plots

```{r S1S2_gc_phylum_plot, fig.height=8, fig.width=10, warning=F}
ggplot(ds, aes(x = S1+S2, y = gc, size = length, color = phylum)) + 
  scale_x_log10() + 
  geom_point(alpha=0.1, color = 'black') +
  geom_point(data=subset(ds, phylum != "NA"), shape = 1, alpha = 0.7) +
  scale_size_area(name= "Scaffold length", max_size=20) +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size = 5, shape = 19)))
```
