Genome extraction
=======================================

## Load data

```{r Loaddata, message=FALSE, warning=FALSE, results='hide'}
library("mmgenome")
knit('Load_data.Rmd', tangle=TRUE)
source(file="Load_data.R")
```

## Initial extraction
The initial genome selection.
```{r zoomA, fig.width=10, fig.height=8}
x <- "S1"
y <- "S2"
data <- d

p <- ggplot(data, aes_string(x = x, y = y, size = "length", color = "phylum")) + 
  scale_x_log10() + 
  scale_y_log10() + 
  geom_point(alpha=0.1, color = 'black') +
  geom_point(data=subset(data, phylum != "NA"), shape = 1, alpha = 0.7) +
  scale_size_area(name= "Scaffold length", max_size=20) +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size = 5, shape = 19)))

#p
#sel <- ggplot_locator(p)

sel <- data.frame(S1  =  c(14.6754, 35.0108, 80.6126, 51.7301, 16.6163, 12.5092),
                  S2  =  c(16.6438, 35.8536, 31.4102, 13.4684, 6.089, 7.5246))

p + 
  geom_polygon(data=sel, aes_string(x, y), size = .5, color = "darkred", fill = NA, linetype= 2) +
  geom_point(data=sel, aes_string(x, y), size = 2, color = "black")
```

## Extract scaffolds and essential genes

Extract all scaffolds and information on essential genes within the defined subspace using the `extract` function.

```{r extractA}
dA<-extract(d, sel)
eA<-extract(e, sel)
```
We store the extracted scaffolds in the variable `dA` and the extracted essential genes in `eA`. `dA` simply stands for `extraction A` of `dataframe d`.

## Calculate statistics on the extracted scaffolds
Then we use the `calc.genome.stats` function to see the basic statistics of the extracted scaffolds.

```{r statsA}
sstats(dA, eA)
```

Look for duplicated single copy genes.

```{r dubA}
head(eduplicates(eA))
```

## First cleanup

We use the pairs function to plot the first 3 components, GC and coverage. 
```{r PCA_pairs_B, fig.width=10, fig.height=10}
data <- dA
palette(gc.trans)

pairs(data[,c("PC1","PC2","PC3", "gc", "S1", "S2")], upper.panel=NULL, col = data$gc-min(d$gc), cex = sqrt(data$length)/75, pch=20)
```

PC2 and PC3 seem to be nice for further subsetting.

```{r zoomB, fig.width=10, fig.height=8}
x <- "PC2"
y <- "PC3"
data <- dA

p <- ggplot(data, aes_string(x = x, y = y, size = "length", color = "phylum")) + 
  #scale_x_log10() + 
  #scale_y_log10() + 
  geom_point(alpha=0.1, color = 'black') +
  geom_point(data=subset(data, phylum != "NA"), shape = 1, alpha = 0.7) +
  scale_size_area(name= "Scaffold length", max_size=20) +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size = 5, shape = 19)))

#p
#sel <- ggplot_locator(p)

sel <- data.frame(PC2  =  c(0.0313, 0.142, 0.264, 0.2705, 0.1241, 0.0232),
                  PC3  =  c(0.022, 0.1095, 0.0708, -0.156, -0.245, -0.1172))

p + 
  geom_polygon(data=sel, aes_string(x, y), size = .5, color = "darkred", fill = NA, linetype= 2) +
  geom_point(data=sel, aes_string(x, y), size = 2, color = "black")
```

```{r extractB}
dB<-extract(dA, sel)
eB<-extract(eA, sel)
```

```{r statsB}
sstats(dB, eB)
```

```{r dubB}
eduplicates(eB)
```

## Using PE connections

We can use the paried-end connections to include repeats and scaffolds missed in the initial extractions as well as obvious contamination. 
```{r network_complete_B}
subset <- extract_nodes(d, dB)
sub.graph <- induced.subgraph(g, subset$nodes)
```

Next we can plot all scaffolds connected to 
```{r plot_network_B}
palette(gc.trans)
plot(sub.graph, layout=layout.fruchterman.reingold)
```

It dosn't look like much in this example. But we do get a few scaffolds pulled in, but also some additional contamination.
```{r subsetC}
dC <- d[subset$scaffolds,]
eC <- e[e$scaffold %in% dC$scaffold,]
```

Looking at the stats.
```{r statsC}
sstats(dC, eC)
```

## Final clean up

We use the pairs function to plot the first 3 components, GC and coverage. 
```{r PCA_pairs_C, fig.width=10, fig.height=10}
data <- dC
palette(gc.trans)

pairs(data[,c("PC1","PC2","PC3", "gc", "S1", "S2")], upper.panel=NULL, col = data$gc-min(d$gc), cex = sqrt(data$length)/75, pch=20)
```

PC1 and PC2 seem to be nice for further subsetting.

```{r zoomC, fig.width=10, fig.height=8}
x <- "PC1"
y <- "PC2"
data <- dC

p <- ggplot(data, aes_string(x = x, y = y, size = "length", color = "phylum")) + 
  #scale_x_log10() + 
  #scale_y_log10() + 
  geom_point(alpha=0.1, color = "black") +
  geom_point(data=subset(data, phylum != "NA"), shape = 1, alpha = 0.7) +
  scale_size_area(name= "Scaffold length", max_size=20) +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size = 5, shape = 19)))

#p
#sel <- ggplot_locator(p)

sel <- data.frame(PC1  =  c(-0.1619, -0.0733, 0.0838, 0.1482, 0.1386, -0.0419, -0.1644),
                  PC2  =  c(0.1572, 0.2698, 0.2747, 0.1808, 0.0779, 0.0015, 0.0098))

p + 
  geom_polygon(data=sel, aes_string(x, y), size = .5, color = "darkred", fill = NA, linetype= 2) +
  geom_point(data=sel, aes_string(x, y), size = 2, color = "black")
```

```{r extractD}
dD<-extract(dC, sel)
eD<-extract(eC, sel)
```

```{r statsD}
sstats(dD, eD)
```

```{r dubD}
eduplicates(eD)
```

Write the scaffolds to a file.
```{r write_sequences}
writeXStringSet(assembly[dD$scaffold], file = "genome1.fa")
```