Stats on essential genes in reference genomes
========================================================

Load needed packages.
```{r Load_packages, message=FALSE, warning=FALSE, results='hide'}
library("mmgenome")
```

Load all the reference data. The reference data was generated by downloading all complete genomes from NCBI, mergeing all genomes and using meta-prodigal for predicting genes. Afterwards HMMER was used to search the essential HMM models against the proteins.
```{r load data}
data(eg)
colnames(eg)
```

## Phylum level

Generate phylum level data.
```{r gen_phylum}
p <- ephylum(eg)
```

Plot the phylum level data. y-legend is Phylum name (Number of genomes, Unique HMMs, Total HMMs).
```{r plot_p, fig.width=12, fig.height=10}
p$plot
```

Look at summary stats on phylum level.
```{r p_summary}
p$table
```

## Species level
Generate species level data. Its highly recommended to subset to a few species first..
```{r gen_species}
eg_alpha <- subset(eg, Class == "Alphaproteobacteria")
s <- especies(eg_alpha)
```

Plot species level data.
```{r plot_species, fig.width=12, fig.height=10}
s$plot
```

Look at summary data.
```{r species_table}
s$table[1:10,c(6,121,122)]
```